A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities

### 2012-03-06 papers/contact/paper.tex
update text and implementation on Yu and Wu's ghs Tested everything on bingley.

\section{Contact density at a given hard
  sphere}\label{contact-at-sphere}

@@

@@

\section{Contact density at a given contact position}

An alternate definition of the inhomogeneous contact density would be
the contact density for spheres \emph{touching at a given point}.
This would be a symmetric measure, since both spheres involved are
treated in the same way.  In fact, the density of such spheres is just
the density $n_0(\xx)$ from FMT.  See Figure~\ref{fig:n0} for an
illustration $n_0$.  Since $n_0$ is the fundamental measure of sphere
surface contact, it would make sense to define a contact density that
would correspond.  This is what is done by Yu and
Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, as discussed in
Section~\ref{sec:yuwu}.

\section{Gross' functional}\label{sec:gross}
We compare with the functional of Gross \emph{et
  al}\cite{gross2009density}, which is of the second kind ($g^b$), where the
contact density is evaluated at the sphere location.
\begin{align}
  n_\textit{contact} &= n_{A} \frac{1 - \frac12 n_{3'}}{\left(1 -
    n_{3'}\right)^3} \\
  n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
\end{align}
where $n_A$ is defined in the introduction and $n_{3'}$ is an averaged
filling fraction defined in a manner analogous to the $n_3$ of FMT,
but with the hard-sphere radius replaced by the hard-sphere diameter,
so the integrals range over spheres that are twice as large.

@@

Yu and Wu developed a functional that should correspond to the contact
density at a given contact
position~$g^a$~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We will
compare both their functional as well as ours with simulation data.

The functional of Yu and Wu for $g_\sigma$ is a symmetric formulation,
but instead of using $n_0$ as the associated density, it uses a
density given by
\begin{align}
  n_{YW} &= n_0 \zeta \\
  \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
\end{align}
where the function $\zeta$ is a measure of local inhomogeneity at the
point of contact.  Given this definition of density, $g_\sigma$ is
given by:
\begin{align}
  g_\sigma^{YW} &= \frac{1}{1-n_3}
    + \frac32 \frac{R n_2}{3}\frac{\zeta}{(1-n_3)^2}
    + \frac12 \frac{\left(\frac{R n_2}{3}\right)^2 \zeta}{(1-n_3)^3}
    \\
  g_\sigma^{YW} &= \frac{1}{1-n_3}
    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}
\end{align}
We should write a bit more about this functional.  I think we can view
it as a third alternative, which we could also predict:
\begin{align}
  g_\sigma^{S'}(\rr) &= \frac1{n_{S'}(\rr)^2}\frac1{4\pi R^2}\int n^{(2)}(\rr - \rr', \rr + \rr')\delta(|\rr'| -
  R)d\rr'
\end{align}
\begin{align}
  n_{S'}(\rr) &= \zeta(\rr) \frac{1}{4\pi R^2}\int n(\rr')\delta(|\rr-\rr'|-R) d\rr'
\end{align}
This is closely related to $g_\sigma^{S}$, but is ever so slightly
different.  In order to be completely fair, we should see about
computing this from our MC calculations.  Unfortunately, it will
require that we compute $\mathbf{n}_2 \cdot \mathbf{n}_2$ which will
be a bit of a pain.  $\ddot\frown$

>>>>

\section{Asymmetrically averaged correlation function}\label{contact-at-sphere}

@@

\section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
We compare with the averaged correlation function of
Gross\cite{gross2009density}, which is of the asymmetrical variety
($g_\sigma^A$), where the correlation function is averaged at location
of one of the spheres in concact.
\begin{align}
  g_\sigma^\text{Gross,A} &= \frac{1 - \frac12 n_{3'}}{\left(1 -
    n_{3'}\right)^3} \\
  n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
\end{align}
where $n_{3'}$ is an averaged
filling fraction defined in a manner analogous to the $n_3$ of FMT,
but with the hard-sphere radius replaced by the hard-sphere diameter,
so the integrals range over spheres that are twice as large.

@@

\section{Symmetrically averaged correlation function}

An alternate definition of the averaged correlation function for an
inhomogeneous system would be the correlation function averaged for
spheres \emph{touching at a given point}.  This would be a symmetric
measure, since both spheres involved are treated in the same way.  In
fact, the density of such spheres is just the density $n_0(\xx)$ from
FMT.  See Figure~\ref{fig:n0} for an illustration $n_0$.  Since $n_0$
is the fundamental measure of sphere surface contact, it makes sense
to define a corresponding correlation function $g_\sigma^S$.  This is
\emph{almost} what is done by Yu and
Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, as discussed in
Section~\ref{sec:yuwu}, although they use a modified density $\zeta
n_0$ to account for density gradients.

@@

Yu and Wu developed a functional that should correspond to the contact
density at a given contact
position~$g^a$~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We
will compare both their functional as well as ours with simulation
data.  Yu and Wu use a symmetric formulation for $g_\sigma$, but
instead of using $n_0$ as the associated density, it uses a density
given by
\begin{align}
  n^\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \\
  \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
\end{align}
where the function $\zeta$ is a measure of local inhomogeneity at the
point of contact.  Given this definition of density, their $g_\sigma$ is
given by:
\begin{align}
  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
    + \frac32 \frac{R n_2}{3}\frac{\zeta}{(1-n_3)^2}
    + \frac12 \frac{\left(\frac{R n_2}{3}\right)^2 \zeta}{(1-n_3)^3}
    \\
  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}
    \\
  g_\sigma^\text{Yu,S} &= \frac{\frac{1}{1-n_3}
    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}}{\zeta^2}
\end{align}
I think the adjusted functional $g_\sigma^{\text{Yu,}S}$ can be directly
compared with our $g_\sigma^S$.  It is implemented in our code.
